Reproducible Research - Project 1
===============================
```{r echo=FALSE}
library(knitr)
opts_chunk$set(cache=T, echo=T, warning=F, message=F)
```

##Read and transform data
Import library "ggplot2"
```{r}
library(ggplot2)
```

Define data type "actDate" to transform it into the data type Date according to the input date format
```{r}
setAs("character","actDate", function(from) as.Date(from, format="%Y-%m-%d") )
```

Download file from the internet and assign to variable df
```{r}
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
df <- read.csv(unz(temp, "activity.csv"), colClasses=c("numeric", "actDate", "numeric"))
unlink(temp)
```

Aggregate data by date over average number of steps
```{r}
aggdf <- aggregate(x=df$steps, by=list(Group.date = df$date), FUN=mean, na.rm=TRUE)
```

##Daily steps by date
```{r}
qplot(aggdf$x, geom="histogram"
      , binwidth=5
      , main="Histogram of daily steps"
      , xlab="Average daily steps"
      , fill=I("white")
      , col=I("black") 
      , alpha=I(.2)
      )
```

Mean: 
```{r, echo=FALSE} 
mean(aggdf$x, na.rm=TRUE)
```

Median:
```{r, echo=FALSE} 
median(aggdf$x, na.rm=TRUE)
```

##Average daily activity pattern
Aggregate data by interval over average number of steps
```{r}
aggdf2 <- aggregate(x=df$steps, by=list(Group.interval = df$interval) , FUN=mean, na.rm=TRUE)
```

Plot of average daily steps by time interval

```{r, echo=FALSE} 
qplot(aggdf2$Group.interval, aggdf2$x, geom="line"
      , main="Daily average steps by time interval"
      , xlab="Time interval"
      , ylab="Average steps")
```

##Time interval that contains the maximum number of steps taken
```{r}
aggdf2[aggdf2$x == max(aggdf2$x), "Group.interval"]
```

##Strategy for filling missing values and new histogram
Calculate all means by time interval
```{r}
mean_by_interval <- aggregate(x=df$steps, by=list(Group.interval = df$interval), FUN=mean, na.rm=TRUE)
```

Fill all missing values using means by interval
```{r}
for (i in 1:nrow(df)){
  if (is.na(df[i, "steps"]))
    df[i, "steps"] <- mean_by_interval[mean_by_interval$Group.interval == df[i, "interval"], "x"]
}
```

Aggregate data by date over average number of steps
```{r}
aggdf <-aggregate(x=df$steps, by=list(Group.date = df$date), FUN=mean, na.rm=TRUE)
```

New histogram of average daily steps
```{r}
qplot(aggdf$x, geom="histogram"
      , binwidth=5
      , main="Histogram of daily steps"
      , xlab="Average daily steps"
      , fill=I("white")
      , col=I("black") 
      , alpha=I(.2)
      )
```

Mean: 
```{r, echo=FALSE} 
mean(aggdf$x, na.rm=TRUE)
```

Median:
```{r, echo=FALSE} 
median(aggdf$x, na.rm=TRUE)
```

##Differences in activity patterns between weekdays and weekends
New factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
df[grepl("Sat|Sun", weekdays(df$date)), "date_type"] <- "weekend"
df[!grepl("Sat|Sun", weekdays(df$date)), "date_type"] <- "weekday"
```

Aggregate data by interval over average number of steps
```{r}
aggdf4 <- aggregate(x=df$steps
                    , by=list(Group.interval = df$interval, Group.date_type = df$date_type)
                    , FUN=mean, na.rm=TRUE)
```

Panel plot containing a time series plot of the time interval and the average number of steps taken averaged across all weekday days or weekend days

```{r}
ggplot(data=aggdf4
       , aes(x=Group.interval, y=x)) +
  ggtitle("Activity pattern between weekdays and weekends") +
  xlab("Time interval") +
  ylab("Average steps") +
  geom_line() + 
  facet_grid(.~ Group.date_type)
```